---
title: "Business Intelligence Lab Submission Markdown"
author: "Lumin"
date: "06/11/23"
output:
  github_document:
    toc: yes
    toc_depth: 4
    fig_width: 6
    fig_height: 4
    df_print: default
always_allow_html: true
editor_options:
  chunk_output_type: console
---


# Student Details

+---------------------------------------------------+------------------------------------------------------------------------------------------------------+
| **Student ID Numbers and Names of Group Members** | 
|                                                   |                                                                                                      |
|                                                   | 1.  112827 - A - Mungai Kenneth                                                                      |
|                                                   |                                                                                                      |
|                                                   | 2.  123324 - B - Kelly Noella Sota                                                                     |
|                                                   |                                                                                                      |
|                                                   | 3.  134265 - A - Emmanuel Kiptoo                                                                   |
|                                                   |                                                                                                      |
|                                                   |                                                                               |
|                                                   |                                                                                                      |
|                                                   |                                                                               |
+---------------------------------------------------+------------------------------------------------------------------------------------------------------+
| **GitHub Classroom Group Name**                   |    Lumin                                                                                                  |
+---------------------------------------------------+------------------------------------------------------------------------------------------------------+
| **Course Code**                                   | BBT4206                                                                                              |
+---------------------------------------------------+------------------------------------------------------------------------------------------------------+
| **Course Name**                                   | Business Intelligence II                                                                             |
+---------------------------------------------------+------------------------------------------------------------------------------------------------------+
| **Program**                                       | Bachelor of Business Information Technology                                                          |
+---------------------------------------------------+------------------------------------------------------------------------------------------------------+
| **Semester Duration**                             | 21^st^ August 2023 to 28^th^ November 2023                                                           |
+---------------------------------------------------+------------------------------------------------------------------------------------------------------+

# Setup Chunk

**Note:** the following "*KnitR*" options have been set as the defaults:\
`knitr::opts_chunk$set(echo = TRUE, warning = FALSE, eval = TRUE, collapse = FALSE, tidy.opts = list(width.cutoff = 80), tidy = TRUE)`.

More KnitR options are documented here <https://bookdown.org/yihui/rmarkdown-cookbook/chunk-options.html> and here <https://yihui.org/knitr/options/>.

```{r setup, include=FALSE}
if (require("formatR")) {
  require("formatR")
} else {
  install.packages("formatR", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}
library(formatR)
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, eval = TRUE,
                      collapse = FALSE, tidy = TRUE)
```

**Note:** the following "*R Markdown*" options have been set as the defaults:

> output:\
> \
> github_document:\
> toc: yes\
> toc_depth: 4\
> fig_width: 6\
> fig_height: 4\
> df_print: default\
> \
> editor_options:\
> chunk_output_type: console


# 1: Install and Load Required Packages

In this step, we ensure that the necessary R packages are installed and loaded. Packages are collections of R functions, data, and compiled code that extend the functionality of R. The install.packages() function is used to install packages, and library() is used to load them.

```{r setup-chunk-one}
## mlbench ----
if (require("mlbench")) {
  require("mlbench")
} else {
  install.packages("mlbench", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

## caret ----
if (require("caret")) {
  require("caret")
} else {
  install.packages("caret", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

## kernlab ----
if (require("kernlab")) {
  require("kernlab")
} else {
  install.packages("kernlab", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

## randomForest ----
if (require("randomForest")) {
  require("randomForest")
} else {
  install.packages("randomForest", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

## naniar ----
if (require("naniar")) {
  require("naniar")
} else {
  install.packages("naniar", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}


```

# 2. Load the Dataset

This code snippet is used to load the dataset "Glass" contained in the mlbench package.

```{r step-two-chunk}
data("Glass")

```

# 3.  Check for Missing Data and Address it

This code checks if there are any missing values.

```{r step-three-chunk}

# Are there missing values in the dataset?
any_na(Glass)

```

# 4.The Resamples Function 

This code trains five machine learning models (LDA, CART, KNN, SVM, and Random Forest) on the "Glass" dataset. It uses cross-validation and collects performance metrics like accuracy and RMSE. The goal is to compare the models' performance.

```{r step-four-chunk}
## 4.a. Train the Models ----
# We train the following models, all of which are using 10-fold repeated cross
# validation with 3 repeats:
#   LDA
#   CART
#   KNN
#   SVM
#   Random Forest


train_control <- trainControl(method = "repeatedcv", number = 10, repeats = 3)


### LDA ----
set.seed(7)
GlassModel_lda <- train(Type ~ ., data = Glass,
                               method = "lda", trControl = train_control)


### CART ----
set.seed(7)
GlassModel_cart <- train(Type ~ ., data = Glass,
                                method = "rpart", trControl = train_control)

### KNN ----
set.seed(7)
GlassModel_knn <- train(Type ~ ., data = Glass,
                               method = "knn", trControl = train_control)

### SVM ----
set.seed(7)
GlassModel_svm <- train(Type ~ ., data = Glass,
                               method = "svmRadial", trControl = train_control)


### Random Forest ----
set.seed(7)
GlassModel_rf <- train(Type ~ ., data = Glass,
                              method = "rf", trControl = train_control)


## 4.b. Call the `resamples` Function ----
# We then create a list of the model results and pass the list as an argument
# to the `resamples` function.

results <- resamples(list(LDA = GlassModel_lda, 
                          CART = GlassModel_cart,
                          KNN = GlassModel_knn, 
                          SVM = GlassModel_svm,
                          RF = GlassModel_rf))

```


# 5. Display the Results

This code does different analyses and visualizations to evaluate machine learning models. It includes a summary table, plots, and statistical tests to compare model performance.

```{r step-five-chunk}
## 1. Table Summary ----
# This is the simplest comparison. It creates a table with one model per row
# and its corresponding evaluation metrics displayed per column.

summary(results)

## 2. Box and Whisker Plot ----
# This is useful for visually observing the spread of the estimated accuracies
# for different algorithms and how they relate.

scales <- list(x = list(relation = "free"), y = list(relation = "free"))
bwplot(results, scales = scales)

## 3. Dot Plots ----
# They show both the mean estimated accuracy as well as the 95% confidence
# interval.

scales <- list(x = list(relation = "free"), y = list(relation = "free"))
dotplot(results, scales = scales)

## 4. Scatter Plot Matrix ----
# This is useful when considering whether the predictions from two
# different algorithms are correlated. If weakly correlated, then they are good
# candidates for being combined in an ensemble prediction.

splom(results)

## 5. Pairwise xyPlots ----
# You can zoom in on one pairwise comparison of the accuracy of trial-folds for
# two models using an xyplot.

# xyplot plots to compare models
xyplot(results, models = c("LDA", "SVM"))

# or
# xyplot plots to compare models
xyplot(results, models = c("SVM", "CART"))


## 6. Statistical Significance Tests ----
# This is used to calculate the significance of the differences between the
# metric distributions of the various models.

### Upper Diagonal ----
# The upper diagonal of the table shows the estimated difference between the
# distributions. If we think that LDA is the most accurate model from looking
# at the previous graphs, we can get an estimate of how much better it is than
# specific other models in terms of absolute accuracy.

### Lower Diagonal ----
# The lower diagonal contains p-values of the null hypothesis.
# The null hypothesis is a claim that "the distributions are the same".
# A lower p-value is better (more significant).

diffs <- diff(results)

summary(diffs)



```